Hierarchical Dispatching and Overriding

### 2017-10-09 NewVersion/sections/Auxiliary.tex
Section 4

\section{Fundamental Notations}\label{subsec:auxdefs}
In this section, we present the fundamental notations used in our formalization.
Some of them are auxiliary functions, others are however important ones as they
directly implement our algorithm for method resolution.

\subsection{Auxiliary functions}
To begin with, we introduce the following basic functions.

@@

\subsection{$\mostSpecific$ and $\mostSpecificOverride$}

@@

\subsection{$\canInstantiate$}
$\canInstantiate(I)$ checks whether interface $I$ can be instantiated by the keyword $\kwnew$.
Below is the algorithm:

@@

\haoyuan{Need more examples?}
\begin{figure*}[t]
	\centering
	\vspace{-1ex}
	\begin{tabular}{ccc}
		\includegraphics[width=1.5cm]{pics/P1.pdf}\hspace{4pt} &
		\includegraphics[width=1.5cm]{pics/P2.pdf}\hspace{4pt} &
		\includegraphics[width=1.5cm]{pics/P3.pdf}\hspace{4pt} \\
		(a) $\mbody(m,C,A) = (A,...)$\ \ \  & (b) $\mbody(m,C,A) = (C,...)$\ \ \  & (c) $\mbody(m,C,A) = (C,...)$
	\end{tabular} \\
	\begin{tabular}{cc}
		\includegraphics[height=3cm]{pics/P4.pdf}\hspace{4pt} &
		\includegraphics[height=3cm]{pics/P5.pdf}\hspace{4pt} \\ 
		(d) $\mbody(m,D,A) = \keyword{undefined}$\ \ \  & (e) $\mbody(m,D,A) = \keyword{undefined}$
	\end{tabular}
	\caption{Examples on $\mbody$. ``$\uparrow$'' stands for hierarchical overriding.}\label{fig:examplesmbody}
	%\saveSpaceFig
\end{figure*}

>>>>

\section{Fundamental Notions}\label{subsec:auxdefs}
In this section, we present the fundamental notions used in our formalization.
Some of them are auxiliary functions, others are however important ones as they
directly implement our algorithm for method resolution.
\bruno{I felt like there's not enough explanation/context in this
  section. It is just dumping definitions without connecting them and
  their purpose properly.}

\subsection{Auxiliary functions}
To begin with, we introduce the following basic functions. 
\bruno{explain: where are these definitions used and for what purpose?
  You need similar explanations everywhere you introduce new definitions.}

@@

\bruno{do not use the names of the definitions as titles of
  Sections. Please give a title that states the purpose. 
For this section in particular it may be something like ``Finding the most
specific method override''}
\subsection{$\mostSpecific$ and $\mostSpecificOverride$}

@@

\subsection{$\canInstantiate$}
$\canInstantiate(I)$ checks whether interface $I$ can be instantiated by the keyword $\kwnew$.
Below is the algorithm:\bruno{used for what and where?}

@@

\haoyuan{Need more examples?}\bruno{ask Marco.}
\begin{figure*}[t]
	\centering
	\vspace{-1ex}
	\begin{tabular}{ccc}
		\includegraphics[width=1.5cm]{pics/P1.pdf}\hspace{4pt} &
		\includegraphics[width=1.5cm]{pics/P2.pdf}\hspace{4pt} &
		\includegraphics[width=1.5cm]{pics/P3.pdf}\hspace{4pt} \\
		(a) $\mbody(m,C,A) = (A,...)$\ \ \  & (b) $\mbody(m,C,A) = (C,...)$\ \ \  & (c) $\mbody(m,C,A) = (C,...)$
	\end{tabular} \\
	\begin{tabular}{cc}
		\includegraphics[height=3cm]{pics/P4.pdf}\hspace{4pt} &
		\includegraphics[height=3cm]{pics/P5.pdf}\hspace{4pt} \\ 
		(d) $\mbody(m,D,A) = \keyword{undefined}$\ \ \  & (e) $\mbody(m,D,A) = \keyword{undefined}$
	\end{tabular}
	\caption{Examples on $\mbody$. ``$\uparrow$'' stands for hierarchical overriding.}\label{fig:examplesmbody}
	%\saveSpaceFig
\end{figure*}

### 2017-10-09 NewVersion/sections/Formalization.tex
Section 4

\section{Formalization}~\label{sec:formalization}
In this section, we present the formal model \MIM{} (\emph{\textbf{F}eatherweight \textbf{H}ierarchical \textbf{J}ava}), based on
Featherweight Java~\cite{Igarashi01FJ}. The syntax, typing rules,
small-step semantics and theorems for type soundness are presented.

@@

The method declaration $\method{I}{m}{I_x}{x}{J}{e}$ introduces a
method named $m$ with result type $I$, parameters $\overline{x}$ of
type $\overline{I_x}$ and the overriding target $J$. The body of the
method is the single statement $ \kwreturn \; e;$. We also allow
defining abstract methods with $\absmethod{I}{m}{I_x}{x}{J}$, which is
similar to a normal method but without a method body. Notably, we have introduced the
\kwoverride{} keyword to override an old implementation of the
method. If the interface that it overrides is exactly the enclosing
interface, then such a method is seen as ``originally defined''; otherwise
it is hierarchical overriding.
For simplicity, overloading is not modelled for methods, which
implies we can uniquely identify a method by its name.\yanlin{the terminology of original methods should have been proposed in Overview.}

@@

\subsection{Subtyping and Typing Rules}
The subtyping of \MIM{} consists of only a few rules shown at the top of Figure~\ref{fig:typingrules}.
In short, subtyping relations are built from the inheritance in interface
declarations. Subtyping is both symmetric, reflexive and transitive.

Details of type-checking rules are displayed at the bottom of Figure~\ref{fig:typingrules}, including expression
typing, well-formedness of methods and interfaces. As a convention, an environment
$\Gamma$ is maintained to store the types of variables, together with
the \kwthis type, namely
the enclosing type. 
% The three rules for method invocation, \textsc{(T-Invk)}, \textsc{(T-PathInvk)} and \textsc{(T-SuperInvk)}
% are very similar, in the sense that they all check the type of the specific method, by using
% an auxiliary function \mtype. \mtype{} is the function for looking up method types, which we will
% illustrate later in Section~\ref{subsec:auxdefs}. After the method
% type is obtained, they all check that the arguments and the receiver
% have compatible types. Additionally, \textsc{(T-PathInvk)} requires the receiver to be the subtype of the specified
% path type, and \textsc{(T-SuperInvk)} checks if the enclosing type directly extends the specified super type.

There are two typing rules for method invocation: \textsc{(T-Invk)} and \textsc{(T-PathInvk)}. \textsc{(T-Invk)} is the typing judgement for a normal method invocation. First it ensures that the method receiver $e_0$ type checks to a type $I_0$. Then it checks the type of the specific method, by using an auxiliary function $\mbody(m, I_0, I_0)$, which is the function for looking up methods that we will illustrate in more detail in Section~\ref{subsec:auxdefs}. The form of \mbody is different to FJ, which has two parameters. The first $I_0$ represents the dynamic type and the second, the static type. If the result of \mbody is not \Undefined, it means the method exists and the corresponding method parameters and return types are given. After the method type is obtained, we check that all arguments $\overline{e}$ type check to $\overline{I}$ which are subtypes of $\overline{J}$. \textsc{(T-PathInvk)} is the typing judgement for a path invocation. Besides the conditions of \textsc{(T-Invk)}, \textsc{(T-PathInvk)} requires the type of receiver to be the subtype of the specified path type. 
% and \textsc{(T-SuperInvk)} checks if the enclosing type directly extends the specified super type.

@@

\subsection{Small-step Semantics and Congruence}
Figure~\ref{fig:smallstep} defines the small-step semantic and
congruence rules, respectively. When evaluating an expression, they
are invoked recursively and alternately to produce a single value in
the end. The small-step semantics rules \textsc{(S-Invk)} and 
\textsc{(S-PathInvk)} behave similarly: each corresponds to one kind of method invocation. \textsc{(S-PathInvk)} invokes \mbody{} for method body lookup. The parameters $\overline{x}$ are replaced with arguments $\overline{v}$ with static type annotations $\overline{\angl{I_x}}$, where $\overline{\angl{I_x}}$ is the parameter types coming from method lookup result. The rule \textsc{(S-Invk)} simply reduces the problem of normal invocation to path invocation.

@@

\subsection{Properties}
Previously the definitions of our model are given, now we proceed to prove the type soundness of 
the model, which relates typing to computation. The type soundness states that, if an expression is 
well-typed, then after many reduction steps it must reduce to a value, with its annotation to be a subtype of the original expression type.
Following the Featherweight Java paper~\cite{Igarashi01FJ}, the type-soundness theorem 
(Theorem~\ref{theorem_soundness}) is proved by using the standard technique of subject reduction (Theorem~\ref{theorem_subject})
and progress (Theorem~\ref{theorem_progress})~\cite{Wright1994}. In Theorem~\ref{theorem_progress} ``$\#(\overline{x})$'' denotes the length of
$\overline{x}$.

@@

\begin{theorem}[Type Soundness]~\label{theorem_soundness}
If $\judgeewf \o {e : I}$ and $e \to^* e'$ with $e'$ a normal form, then $e'$ is 
a value $v$ with $\judgeewf \o {v:I}$.
\end{theorem}
\begin{proof}
Immediate from Theorem~\ref{theorem_subject} and Theroem~\ref{theorem_progress}.
\end{proof}

>>>>

\section{Formalization}~\label{sec:formalization}
In this section, we present the formal model \MIM{} (\emph{\textbf{F}eatherweight \textbf{H}ierarchical \textbf{J}ava}), based on
Featherweight Java~\cite{Igarashi01FJ}. The syntax, typing rules,
small-step semantics and theorems for type soundness are presented.
\bruno{In general, please try to use latex paragraphs to add more
  structure and emphasize contents.}

@@

The method declaration $\method{I}{m}{I_x}{x}{J}{e}$ introduces a
method named $m$ with result type $I$, parameters $\overline{x}$ of
type $\overline{I_x}$ and the overriding target $J$. The body of the
method is the single statement $ \kwreturn \; e;$. We also allow
defining abstract methods with $\absmethod{I}{m}{I_x}{x}{J}$, which is
similar to a normal method but without a method body. Notably, we have introduced the
\kwoverride{} keyword to override an old implementation of the
method. If the interface that it overrides is exactly the enclosing
interface, then such a method is seen as ``originally defined''; otherwise
it is hierarchical overriding.
For simplicity, overloading is not modelled for methods, which
implies that we can uniquely identify a method by its name.\yanlin{the terminology of original methods should have been proposed in Overview.}
\bruno{Then please do so.}

@@

\subsection{Subtyping and Typing Rules}
\paragraph{Subtyping}
The subtyping of \MIM{} consists of only a few rules shown at the top of Figure~\ref{fig:typingrules}.
In short, subtyping relations are built from the inheritance in interface
declarations. Subtyping is both symmetric, reflexive and transitive.

\paragraph{Type-checking}
Details of type-checking rules are displayed at the bottom of Figure~\ref{fig:typingrules}, including expression
typing, well-formedness of methods and interfaces. As a convention, an environment
$\Gamma$ is maintained to store the types of variables, together with
the \kwthis type, namely
the enclosing type. 
% The three rules for method invocation, \textsc{(T-Invk)}, \textsc{(T-PathInvk)} and \textsc{(T-SuperInvk)}
% are very similar, in the sense that they all check the type of the specific method, by using
% an auxiliary function \mtype. \mtype{} is the function for looking up method types, which we will
% illustrate later in Section~\ref{subsec:auxdefs}. After the method
% type is obtained, they all check that the arguments and the receiver
% have compatible types. Additionally, \textsc{(T-PathInvk)} requires the receiver to be the subtype of the specified
% path type, and \textsc{(T-SuperInvk)} checks if the enclosing type directly extends the specified super type.

There are two typing rules for method invocation: \textsc{(T-Invk)}
and \textsc{(T-PathInvk)}. \textsc{(T-Invk)} is the typing judgement
for a normal method invocation. First it ensures that the method
receiver $e_0$ type checks to a type $I_0$. Then it checks the type of
the specific method, by using an auxiliary function $\mbody(m, I_0,
I_0)$, which is the function for looking up methods that we will
illustrate in more detail in Section~\ref{subsec:auxdefs}. The form of
\mbody is different to FJ\bruno{Use xspace to fix the spacing problem
  of macros like mbody}, which has two parameters. The first $I_0$
represents the dynamic type and the second, the static type. If the
result of \mbody is not \Undefined, it means that the method exists and the corresponding method parameters and return types are given. After the method type is obtained, we check that all arguments $\overline{e}$ type check to $\overline{I}$ which are subtypes of $\overline{J}$. \textsc{(T-PathInvk)} is the typing judgement for a path invocation. Besides the conditions of \textsc{(T-Invk)}, \textsc{(T-PathInvk)} requires the type of receiver to be the subtype of the specified path type. 
% and \textsc{(T-SuperInvk)} checks if the enclosing type directly extends the specified super type.

@@

\subsection{Small-step Semantics and Congruence}
Figure~\ref{fig:smallstep} defines the small-step semantic and
congruence rules, respectively. When evaluating an expression, they
are invoked recursively and alternately to produce a single value in
the end. The small-step semantics rules \textsc{(S-Invk)} and 
\textsc{(S-PathInvk)} behave similarly: each corresponds to one kind of method invocation. \textsc{(S-PathInvk)} invokes \mbody{} for method body lookup. The parameters $\overline{x}$ are replaced with arguments $\overline{v}$ with static type annotations $\overline{\angl{I_x}}$, where $\overline{\angl{I_x}}$ are the parameter types coming from method lookup result. The rule \textsc{(S-Invk)} simply reduces the problem of normal invocation to path invocation.

@@

\subsection{Properties}
Now we proceed to prove the type soundness of 
the model, which relates typing to computation. Type soundness states that, if an expression is 
well-typed, then after many reduction steps it must reduce to a value, with its annotation to be a subtype of the original expression type.
Following the Featherweight Java paper~\cite{Igarashi01FJ}, the type-soundness theorem 
(Theorem~\ref{theorem_soundness}) is proved by using the standard technique of subject reduction (Theorem~\ref{theorem_subject})
and progress (Theorem~\ref{theorem_progress})~\cite{Wright1994}. In Theorem~\ref{theorem_progress} ``$\#(\overline{x})$'' denotes the length of
$\overline{x}$.

@@

\begin{theorem}[Type Soundness]~\label{theorem_soundness}
If $\judgeewf \o {e : I}$ and $e \to^* e'$ with $e'$ a normal form, then $e'$ is 
a value $v$ with $\judgeewf \o {v:I}$.
\end{theorem}
\begin{proof}
Immediate from Theorem~\ref{theorem_subject} and Theroem~\ref{theorem_progress}.
\end{proof}
\bruno{Were we not supposed to have some note/comment as to why the
  type soundness theorem is stricter that FJ?}

### 2017-10-09 NewVersion/sections/Introduction.tex
Section 4

C++ and C\# do allow for two
unintentionally conflicting methods to coexist in a class. 
C++ accepts the triangle inheritance and
resolves the ambiguity by \emph{static dispatching} or casts. But the
drawback is that these conflicting methods cannot be further overriden. 
\bruno{how about C\#?}
% However, C++ has
% limited support for virtual methods with unintentional conflicts, and
% it will often throw errors when composing them. This is again
% unsatisfactory because virtual methods are pervasive in OOP and used 
% for code reuse and extensibility. A problem with the C++ approach is
% that programmers can only use either static or dynamic dispatching separately, but dealing
% with unintentional method conflicts seems to require a combination of
% both. 
Some other workarounds or approaches include delegation and
renaming/exclusion in the trait model. However renaming/exclusion 
can break the subtyping relation between a subclass and its parent.
This is not adequate for the class model commonly used in mainstream 
OOP languages, where the subclass is always expected to be a subtype 
of the parent class.

@@

This paper proposes two mechanisms to deal with unintentional method
conflicts: \textit{hierarchical dispatching} and \emph{hierarchical
  overriding}. Hierarchical dispatching is inspired by the mechanisms in C++ and
provides an approach
to method dispatching, which combines static and dynamic
information. Using hierarchical dispatching, the method binder will look
at both the \emph{static type} and the \emph{dynamic type} of the
receiver during runtime. When there are multiple branches that cause
unintentional conflicts, the static type can specify one branch among
them for unambiguity, and the dynamic type helps to find the most
specific implementation. In that case, both unambiguity and
extensibility are preserved. 
\textit{Hierarchical overriding} allows method overriding to be applied
only to one branch of the class hierarchy. Hierarchical overriding 
adds expressive power that is not available in languages such as 
C++ or C\#. To present both ideas, we introduce a
formalized model \MIM{} in Section~\ref{sec:formalization} based on
Featherweight Java~\cite{Igarashi01FJ}, together with theorems and
proofs for type soundness.

>>>>

C++ and C\# do allow for two
unintentionally conflicting methods to coexist in a class. 
C++ accepts the triangle inheritance and
resolves the ambiguity by \emph{static dispatching} or casts. But the
drawback is that these conflicting methods cannot be further overriden. 
\bruno{how about C\#? Say something about C\# here.}
% However, C++ has
% limited support for virtual methods with unintentional conflicts, and
% it will often throw errors when composing them. This is again
% unsatisfactory because virtual methods are pervasive in OOP and used 
% for code reuse and extensibility. A problem with the C++ approach is
% that programmers can only use either static or dynamic dispatching separately, but dealing
% with unintentional method conflicts seems to require a combination of
% both. 
Some other workarounds or approaches include delegation and
renaming/exclusion in the trait model. However renaming/exclusion 
can break the subtyping relation between a subclass and its parent.
This is not adequate for the class model commonly used in mainstream 
OOP languages, where the subclass is always expected to be a subtype 
of the parent class.

@@

This paper proposes two mechanisms to deal with unintentional method
conflicts: \textit{hierarchical dispatching} and \emph{hierarchical
  overriding}. Hierarchical dispatching is inspired by the mechanisms in C++ and
provides an approach
to method dispatching, which combines static and dynamic
information. Using hierarchical dispatching, the method binder will look
at both the \emph{static type} and the \emph{dynamic type} of the
receiver during runtime. When there are multiple branches that cause
unintentional conflicts, the static type can specify one branch among
them for unambiguity, and the dynamic type helps to find the most
specific implementation. In that case, both unambiguity and
extensibility are preserved. The main novelty over existing work is 
the formalization of the essence of a hierarchical dispatching
algorithm, which (as far as we know) has not been formalized before. 
Furthermore, \textit{hierarchical overriding} allows method overriding to be applied
only to one branch of the class hierarchy. Hierarchical overriding 
adds expressive power that is not available in languages such as 
C++ or C\#. To present both ideas, we introduce a
formalized model \MIM{} in Section~\ref{sec:formalization} based on
Featherweight Java~\cite{Igarashi01FJ}, together with theorems and
proofs for type soundness.
